---
title: "Tracking Delays: Analysing the Impact Precipitation on Swiss Railway Punctuality"
author: "Thomas Jucker"
date: "18 December 2025"
bibliography: references.bib  
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

delayed on arrival only! because delayed at departure might be due to other factors, e.g. passengers


```{r setup, include=FALSE}
# global options for "knitting"
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(readr)
library(ggplot2)
library(gganimate)
library(gifski)
library(here)
library(readr)

source(here("R/analyse_modobs2.R"))
```


# Introduction
Gives the background, explains the relevance, and motivates the (research) question. The introduction may also motivate the methodological approach, but doesn’t give its details.

# Methods and Data:
Describes the data (scale, source, etc.) and the methods.

# Results: Reports the results in a logical order that facilitates reading - from the general pattern to the details. Basic and central finding first, then the nuance.
Discussion: Explain what the results mean in relation to the question stated in the Introduction. Discuss caveats, compare to earlier findings, and explain the relevance of your results.


## References
