---
title: "Tracking Delays: Analysing the Impact of Precipitation on Swiss Railway Punctuality"
author: "Thomas Jucker"
date: "18 December 2025"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---


# Introduction
Gives the background, explains the relevance, and motivates the (research) question. The introduction may also motivate the methodological approach, but doesn’t give its details.

Railway operation is affected by a range of environmental factors. In Switzerland,
these effects are accentuated in autumn, when precipitation increases, temperatures
fall and many trees drop their leaves. It is common knowledge that this season is
prone to lower punctuality, which is a measure of quality of this mode of transport. 
This project investigates the influence of one of these variables, precipitation.
The script is built to analyse a recent 24h timeframe.


# Methods and Data:
First, daily public transport data is obtained from an open data source
"(opentransportdata.swiss)". The raw data contains one data point for every ride
conducted on that day. E.g. The stop of train service R16 in Biel with planned
and actual arrival and departure time. Because the dataset also contains 
non-swiss stops and bus and ferry services, its filtered for to the relevant
extent first. Then delay on arrival is calculated from planned vs. actual values.
Delays on departure are ignored because it is assumed that they would not be 
caused by precipitation but rather from human influence. 
Then the train stations are assigned their geographical coordinates by matching
the BPUIC number to an open source table containing their location
"(opentransportdata.swiss)". Next, the data is aggregated to show delay rate 
per stop and per hour. Precipitation data stems from the radar product
CombiPrecip"(CPC)" by the Federal Office of Meteorology and Climatology 
MeteoSwiss. Data is provided as .h5 raster files with a time resolution of 
10 minutes, containing the precipitation total of the previous 60 minutes. 
Processing is conducted to match the hourly punctuality data structure. 
Hourly precipitation total is then extracted at the exact locations of the 
train stations. Eventually, global and hourly correlation is investigated.

## Results

# Run Script

Run the complete workflow to create necessary environment to display 
punctuality, precipitation and correlation between the two.

```{r}
source("R/00_script_consolidation.R", local = TRUE)

```

## Punctuality
```{r}
# Leaflet Map for specific hour
hour_to_plot <- 8   # choose the hour you want to display
delay_hour_subset <- delay_sf_hourly_wgs %>%
  filter(hour == hour_to_plot)

leaflet(delay_hour_subset) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis(100, option = "plasma")[cut(delay_rate, breaks = 100)],
    stroke = FALSE,
    fillOpacity = 0.8,
    popup = ~paste0(
      "<b>", HALTESTELLEN_NAME, "</b><br>",
      "Hour: ", hour, ":00<br>",
      "Delay rate: ", round(delay_rate * 100, 1), " %<br>",
      "Trains delayed: ", n_delayed, " / ", n_trains
    )
  ) %>%
  addLegend(
    position = "bottomright",
    pal = colorNumeric(palette = "plasma", domain = delay_hour_subset$delay_rate),
    values = delay_hour_subset$delay_rate,
    title = paste0("Delay rate @ ", hour_to_plot, ":00")
  )

```

Hourly punctuality usually shows high regional variability. A delay rate of 0 
means no trains were delayed during that hour on that stop. A delay rate of 1 
means all trains were delay. Selecting the stops displays the amount of trains.

## Precipitation 

```{r}
leaflet() %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  addRasterImage(first_hour, colors = pal_precip, opacity = 0.7) %>%
  addLegend(pal = pal_precip, values = values(first_hour),
            title = paste0("Hourly Precipitation [mm] (Hour ", hour_to_plot, ")"))
```

Hourly precipitation total is display can be explored through an interactive
map. This allows for a first visual, spatial comparison of precipitation and 
delay concentration.

## Punctuality and Precipitation Synthesis Map
```{r}
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  
  # Raster layer for precipitation
  addRasterImage(
    precip_r,
    colors = pal_precip,
    opacity = 0.6,
    group = "Precipitation"
  ) %>%
  
  # Delay points layer
  addCircleMarkers(
    data = delay_hour_subset,
    radius = 5,
    color = ~pal_delay(delay_rate),
    stroke = FALSE,
    fillOpacity = 0.9,
    popup = ~paste0(
      "<b>", HALTESTELLEN_NAME, "</b><br>",
      "Hour: ", hour, ":00<br>",
      "Delay rate: ", round(delay_rate * 100, 1), "%<br>",
      "Precipitation: ", round(precip_mm, 2), " mm"
    ),
    group = "Delay rate"
  ) %>%
  
  
  # Legends
  addLegend(
    "bottomright", pal = pal_delay, values = delay_hour_subset$delay_rate,
    title = paste0("Delay rate @ ", hour_to_plot, ":00")
  ) %>%
  addLegend(
    "bottomleft", pal = pal_precip, values = values(precip_r),
    title = paste0("Hourly precipitation (mm) @ ", hour_to_plot, ":00")
  )
```

Merging the two variables into one map yields the precise precipitation
volumes at the stations.

## Correlation

The pearson correlation between hourly precipitation and punctuality shows
no significant relationship. Correlation values are comparatively low and also 
not consistently positive or negative for different hours of the day. 

```{r}
print(cor_hourly, n = Inf)
```
Plotting the variables on a scatterplot confirms this. The different data points,
punctuality and precipitation at a specific stop, seem to show no apparent pattern.

```{r}
ggplot(analysis_df, aes(x = delay_rate, y = precip_mm)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
  labs(
    title = "Correlation between Precipitation and Train Delay Rate",
    x = "Precipitation (mm)",
    y = "Delay rate"
  )
```


# Discussion: 
The results suggest that precipitation is not *the* single, significant factor
influencing train punctuality. However, the results should be considered with
caution. The analysis conducted contains some important shortcomings. Firstly, 
the time period of 24 hours is too short. By analysing a longer time series, 
e.g. 5 years, seasonal patterns might be detected. Secondly, precipitation values
are extracted at the coordinates of the train stops only. Including precipitation
at the immediate sourounding of a stop might improve the model. Thirdly, train delays can
arise from a multitude of influences. Adding more environmental values such as
type of precipitation, temperature or humidity or correcting for human influence
could result in a better r2 score. Fourthly, instead of using single stop
stations, analysing train services along lines (e.g. Zurich - Olten - Bern) 
could also improve significance of the findings. 

